Assessing the spatial patterning of Middle Paleolithic human settlement in Western Iberia

### 2020-11-01 analysis/paper/paper.Rmd
added period to dataset

---
title: "Assessing the spatial patterning of Middle Paleolithic human settlement in Western Iberia"
author:
  - João Cascalheira
  - Célia Gonçalves
  - Daniela Maio
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  Middle Paleolithic; Settlement patterns; GIS; Western Iberia
---

@@

```{r, setup, eval = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

@@

#Import geo tiffs
areaDEM <- raster("../data/raw_data/areaDEM.tif")
areaDEM_cut <- raster("../data/raw_data/areaDEM_cut.tif")
areaSLOPE_cut <- raster("../data/raw_data/areaSLOPE_cut.tif")
areaASPECT_cut <- raster("../data/raw_data/areaASPECT_cut.tif")
areaDIST_cut <- raster("../data/raw_data/areaDIST_cut2.tif")

# Convert coordinates to longlat WGS84
areaDEM <- projectRaster(areaDEM, crs="+proj=longlat +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
areaDEM_cut <- projectRaster(areaDEM_cut, crs="+proj=longlat +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
areaSLOPE_cut <- projectRaster(areaSLOPE_cut, crs="+proj=longlat +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
areaASPECT_cut <- projectRaster(areaASPECT_cut, crs="+proj=longlat +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
areaDIST_cut <- projectRaster(areaDIST_cut, crs="+proj=longlat +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")

#Import sites and convert to shapefile
sites <- read.csv("../data/raw_data/sites_short_list.csv", header=TRUE)
#caves <- read.csv("../data/raw_data/all_caves.csv", header=TRUE)

#Convert to shapefile
coordinates(sites)<-~Longitude+Latitude
proj4string(sites) = CRS("+init=epsg:4326")
writeOGR(sites, "../data/raw_data/spatial_data/shapefiles","sites_short", "ESRI Shapefile")

@@

# Extract variables

# Resample Distance raster
areaDIST_cut <- resample(areaDIST_cut,areaDEM_cut, resample='bilinear')

# Stack maps
terrainstack <- stack(areaDEM_cut,
                      areaSLOPE_cut,
                      areaASPECT_cut,
                      areaDIST_cut)

@@

caves_vals <- raster::extract(terrainstack,
                              caves_long,
                              buffer = 20,
                              fun = mean,
                              sp = TRUE)

@@

general_table <- as.data.frame(tabmulti(data = as.data.frame(sites_vals), xvarname = "Type",
              yvarnames = c("areaDEM_cut", "areaSLOPE_cut", "areaASPECT_cut", "areaDIST_cut2"),
              ymeasures = c("median", "median", "median", "median"), , n.headings = TRUE,
              yvarlabels = list(areaDEM_cut = "Elevation", areaSLOPE_cut = "Slope", areaASPECT_cut = "Aspect", areaDIST_cut2 = "Distance")))

@@

@@

pca_data_active2 <- pca_data %>%  
  drop_na() %>% 
  dplyr::select(Elevation, Slope, Aspect, Distance, Latitude)

@@

groups_2 <- as.factor(pca_data$Type)

@@

dim_1 <- as.data.frame(corr_table$Dim.1$quanti)
round(dim_1, 3)

dim_2 <- as.data.frame(corr_table$Dim.2$quanti)
round(dim_2, 3)

>>>>

---
title: "Assessing the spatial patterning of Middle Paleolithic human settlement in Western Iberia"
author:
  - João Cascalheira
  - Célia Gonçalves
  - Daniela Maio
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/elsevier-harvard.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  Middle Paleolithic; Settlement patterns; GIS; Western Iberia
---

@@

```{r, setup, eval = FALSE, echo = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

@@

#Import geo tiffs
areaDEM <- raster("../data/raw_data/areaDEM_2.tif")
areaDEM_cut <- raster("../data/raw_data/areaDEM_cut.tif")
areaASPECT_cut <- raster("../data/raw_data/areaASPECT_cut.tif")
areaSLOPE_cut <- raster("../data/raw_data/areaSLOPE_cut.tif")
areaDIST_cut <- raster("../data/raw_data/areaDIST_cut2.tif")

# Calculate SLOPE and ASPECT
areaASPECT_cut <- resample(areaASPECT_cut, areaDEM_cut, method = "bilinear")
areaSLOPE_cut <- resample(areaSLOPE_cut, areaDEM_cut, method = "bilinear")

#Import sites and convert to shapefile
#sites <- read.csv("../data/raw_data/sites_short_list.csv", header=TRUE)
#caves <- read.csv("../data/raw_data/all_caves.csv", header=TRUE)

#coordinates(sites)<-~Longitude+Latitude
#proj4string(sites) = CRS("+init=epsg:4326")
#writeOGR(sites, "../data/raw_data/spatial_data/shapefiles","sites_short", "ESRI Shapefile")

@@

# Stack maps
terrainstack <- stack(areaDEM_cut,
                      areaASPECT_cut,
                      areaSLOPE_cut,
                      areaDIST_cut)

@@

caves_vals <- raster::extract(terrainstack,
                              caves,
                              buffer = 20,
                              fun = mean,
                              sp = TRUE)

@@

general_table <- as.data.frame(tabmulti(data = as.data.frame(sites_vals), xvarname = "Type",
              yvarnames = c("areaDEM_cut", "areaSLOPE_cut", "areaASPECT_cut", "areaDIST_cut2"),
              ymeasures = c("median", "median", "median", "median"), n.headings = TRUE,
              yvarlabels = list(areaDEM_cut = "Elevation", areaSLOPE_cut = "Slope", areaASPECT_cut = "Aspect", areaDIST_cut2 = "Distance")))

@@

```{r pca-table, eval = FALSE, echo = FALSE}

pca_data_active2 <- pca_data %>% 
  dplyr::select(Elevation, Slope, Aspect, Distance, Latitude)

dim_1 <- as.data.frame(corr_table$Dim.1$quanti)

dim_1 <- dim_1 %>% 
  rownames_to_column("Variable") %>% 
  mutate(PC = c("PC1", "", "", "", "")) %>% 
  dplyr::select(PC, Variable, correlation, p.value)

dim_2 <- as.data.frame(corr_table$Dim.2$quanti)

dim_2 <- dim_2 %>% 
  rownames_to_column("Variable") %>% 
  mutate(PC = c("PC2", "")) %>% 
  dplyr::select(PC, Variable, correlation, p.value)

pca_table <- rbind(dim_1, dim_2)

knitr::kable(pca_table, caption = "Correlation coefficients between variables and the first two Principal Components, and corresponding significance test result.")

@@

groups <- pca_data %>% 
  dplyr::select(Dmn_R_M)
groups <- as.factor(groups$Dmn_R_M)

@@

@@
